---
title: "Task"
author: "Student Number: 2967794j"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  eval: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: libraries
library(ggplot2)
library(tidyverse)
library(gt)
library(patchwork)
library(gridExtra)
library(moderndive)
```

```{r}
#| label: data
cats <- read.csv("C:/Users/2967794j/Desktop/cats.csv")
```

# Introduction {#sec-Intro}

this is a document

# Exploratory Data Analysis {#sec-EDA}
displays the mean, median and standard deviation for sepal width and length for each of the three different species of iris. From @tbl-summary, we can see that the sepal length is, on average, greater than sepal width. The species *v.* *setosa* has the shortest, on average, sepal length, followed by *v.* *versicolor* and then *v.* *virginica* with the longest average petal length. However, the species *setosa* has the largest, on average, sepal width, followed by *v.* *virginica* and *v.* *versicolor*, respectively. From the standard deviation, the variability in sepal width is relatively similar for each species, however, there is more variability in sepal length, with the standard deviation for the species *v.* *virginica* almost twice that of the species *v.* *setosa*.

```{r}
#| echo: false
#| label: tbl-summary
#| tbl-cap:  Mean, median and standard deviation (sd) min max iqr sample_size of cats.
cats |>
summarize('Mean' = mean(Hwt),
'Median' = median(Hwt),
'St.Dev' = sd(Hwt),
'Min' = min(Hwt),
'Max' = max(Hwt),
'IQR' = quantile(Hwt,0.75)-quantile(Hwt,0.25),
'Sample_size' = n(),
.by = Sex) |>
gt() |>
fmt_number(decimals=2) |>
cols_label(
Mean = html("Mean"),
Median = html("Median"),
St.Dev = html("Std. Dev"),
Min = html("Minimum"),
Max = html("Maximum"),
IQR = html("Interquartile Range"),
Sample_size = html("Sample Size")
)
```

# Formal DataAnalysis {#sec-FDA}
$$\widehat{\mbox{Hwt}} = \widehat{\alpha} +
\widehat{\beta}_{\mbox{Male}} \cdot \mathbb{I}_{\mbox{Male}}(x) $$
where

• the intercept  $\widehat{\alpha}$ is the mean heart weight for the baseline category
of Females;

• $\widehat{\beta}_{\mbox{Male}}$ is the difference in the mean heart weight
of a Males relative to the baseline category Females; and

• 𝕀Male(𝑥) $\mathbb{I}_{\mbox{Male}}(x)$ is an indicator function such that
𝕀Male(𝑥) = { 1 if Sex of 𝑥th observation is Male,
0 Otherwise.
$$\mathbb{I}_{\mbox{Male}}(x)=\left\{
\begin{array}{ll}
1 ~~~ \mbox{if Sex of} ~ x \mbox{th observation is Male},\\
0 ~~~ \mbox{Otherwise}.\\
\end{array}
\right.$$
When this model is fitted to the data, the following estimates of $\widehat{\alpha}$ (intercept) and $\mathbb{I}_{\mbox{Male}}(x)$
(SexM) are returned:
```{r}
model <- lm(Hwt ~ Sex, data = cats)
Coefs <- round(coef(model), 3)
table_values <- get_regression_table(model)
table_values |>
gt() |>
fmt_number(decimals=2) |>
tab_style(style = cell_text(weight = "bold"),
locations = cells_column_labels())
```
# Conclusions {#sec-Conc}
