---
title: "Group work 2"
author: "Group 04"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  eval: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: libraries
library(ggplot2)
library(tidyverse)
library(gt)
library(patchwork)
library(gridExtra)
library(moderndive)
library(GGally)
library(skimr)
```

```{r}
#| label: data
data <- read.csv("C:/Users/2962099Q/Desktop/Project 2/dataset04.csv")
data <- data %>%
  select(-Region)%>%
  relocate(6, .before = everything())
colnames(data) <- c("Family_members","Income","Food_Expenditure","Sex","Head_Age","Household_Type","Floor_Area","House_Age","Num_bedrooms","Electricity")
write.csv(data,"tidy_Data.csv")
```

# Introduction {#sec-Intro}

Understanding the factors that influence household size is crucial for
policymakers in planning housing, utilities, and social programs. This
study analyses data from the Family Income and Expenditure Survey (FIES)
in the Philippines to identify key household-related variables that
impact the number of people living in a household. Using a Generalized
Linear Model (GLM), we will examine how factors such as household
income, food expenditure, house size, number of bedrooms, and household
head characteristics influence family size. The findings from this
analysis can provide valuable insights for government officials to
support effective resource allocation and housing policies across
different regions in the Philippines.

# Exploratory Data Analysis {#sec-EDA}

```{r}
#| echo: false
#| label: tbl-summary
#| tbl-cap:  Mean, median and standard deviation (sd) min max iqr sample size of each factor.
summary_table <- tibble(
  Variable = c("Family_Members", "Income", "Food_Expenditure", "Head_Age", "Floor_Area", "House_Age", "Num_beadrooms"),
  Mean = c(mean(data$Family_members), mean(data$Income), mean(data$Food_Expenditure), mean(data$Head_Age) ,mean(data$Floor_Area) ,mean(data$House_Age), mean(data$Num_bedrooms)),
  Median = c(median(data$Family_members), median(data$Income), median(data$Food_Expenditure), median(data$Head_Age), median(data$Floor_Area) ,median(data$House_Age), median(data$Num_bedrooms)),
  St.Dev = c(sd(data$Family_members), sd(data$Income), sd(data$Food_Expenditure), sd(data$Head_Age) ,sd(data$Floor_Area) ,sd(data$House_Age), sd(data$Num_bedrooms)),
  Min = c(min(data$Family_members), min(data$Income), min(data$Food_Expenditure), min(data$Head_Age),min(data$Floor_Area) ,min(data$House_Age), min(data$Num_bedrooms)),
  Max = c(max(data$Family_members), max(data$Income), max(data$Food_Expenditure), max(data$Head_Age),max(data$Floor_Area) ,max(data$House_Age), max(data$Num_bedrooms)),
  IQR = c(
    quantile(data$Family_members, 0.75) - quantile(data$Family_members, 0.25),
    quantile(data$Income, 0.75) - quantile(data$Income, 0.25),
    quantile(data$Food_Expenditure, 0.75) - quantile(data$Food_Expenditure, 0.25),
    quantile(data$Head_Age, 0.75) - quantile(data$Head_Age, 0.25),
    quantile(data$Floor_Area, 0.75) - quantile(data$Floor_Area, 0.25),
    quantile(data$House_Age, 0.75) - quantile(data$House_Age, 0.25),
    quantile(data$Num_bedrooms, 0.75) - quantile(data$Num_bedrooms, 0.25)
  ),
  Sample_size = c(length(data$Family_members), length(data$Income), length(data$Food_Expenditure), length(data$Head_Age),length(data$Floor_Area),length(data$House_Age),length(data$Num_bedrooms))
)


summary_table |>
  gt() |>
  fmt_number(columns = vars(Mean, Median, St.Dev, Min, Max, IQR, Sample_size), decimals = 2) |>
  cols_label(
    Mean = html("Mean"),
    Median = html("Median"),
    St.Dev = html("Std. Dev"),
    Min = html("Minimum"),
    Max = html("Maximum"),
    IQR = html("Interquartile Range"),
    Sample_size = html("Sample Size")
  )
```

Table 1 presents descriptive statistics for household-related variables.
The number of family members has a mean of 4.53 and a median of 4.00,
with a standard deviation of 2.22, indicating moderate variation in
household size. Income shows a large discrepancy between the mean
(182,984.80) and median (120,362.00), suggesting a skewed distribution,
which may indicate that higher-income households could support larger
families. Floor area has a high standard deviation (34.67), implying
considerable variation in living space, which could influence household
size. The number of bedrooms has a small interquartile range (1.00),
suggesting most households have similar bedroom availability, which may
constrain family size. These factors are key in analyzing what
influences the number of people living in a household.

```{r}
#| echo: false
#| fig-cap: Relationship between family members and household head sex.
#| label: fig-Sex
#| fig-align: center
##| fig-width: 4.5
##| fig-height: 3.5


ggplot(data, aes(x = Sex, y = Family_members, color = Sex)) +
  geom_boxplot() +
  labs(x = "Sex", y = "Family_members", color = "Sex") 
```

Figure 1 shows the relationship between family size and the sex of the
household head. Households led by males tend to have a slightly higher
median number of family members than those led by females. The spread of
family sizes is also wider for male-headed households, with more
outliers at the higher end, indicating that larger families are more
commonly associated with male household heads.

```{r}
#| echo: false
#| fig-cap: Relationship between family members and type of household.
#| label: fig-Household
#| fig-align: center
#| fig-width: 10
##| fig-height: 3.5


ggplot(data, aes(x = Household_Type, y = Family_members, color=Household_Type)) +
  geom_boxplot() +
  labs(x = "Household_Type", y = "Family_members", color = "Household_Type")
```

Figure 2 illustrates the relationship between family size and household
type. Extended families have the highest median number of family
members, followed by households with two or more non-related members.
Single-family households tend to have the smallest family size, with
fewer extreme outliers. This suggests that household structure
significantly influences the number of people living in a household,
with extended families naturally accommodating more members.

```{r}
#| echo: false
#| fig-cap: Histograms of Total Number of Family members.
#| label: fig-Family-members.
#| fig-align: center
#| fig-width: 5
##| fig-height: 3.5

ggplot(data, aes(x = Family_members)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Distribution of Total Number of Family members", x = "Total Number of Family members", y = "Times")
```

Figure 3 illustrates the distribution of the total number of family
members in the dataset. The distribution is right-skewed, with the
majority of households having between 3 and 6 family members. The most
common household size is 4 members, followed by households with 3 and 5
members. Very few households have more than 10 members, and households
with fewer than 3 members are also relatively rare. This suggests that
medium-sized households are the most prevalent in the sample population,
while both very small and very large families are less common.

```{r}
#| echo: false
#| fig-cap: Correlation plot.
#| label: fig-correlationship
#| fig-align: center
#| fig-width: 15
#| fig-height: 15


ggpairs(data[, c("Family_members","Income", "Food_Expenditure",  "Head_Age",  "Floor_Area", "House_Age", "Num_bedrooms", "Electricity")], 
        upper=list(continous=wrap("points",alpha=0.4,color="#d73027")),
        lower="blank",axisLabels="none")
```

Figure 4 presents a correlation plot of household-related variables. The
number of family members has a weak positive correlation with income
(0.124) and food expenditure (0.425), suggesting that higher-income
households may support larger families, and food expenses increase
accordingly. There is also a weak positive correlation between family
size and the number of bedrooms (0.165), indicating that larger
households tend to have more bedrooms. Other variables, such as house
floor area and electricity access, show minimal correlation with family
size. These insights suggest that economic factors and housing capacity
may play a role in determining household size. A Generalized Linear
Model (GLM) will further analyse these relationships.

# Formal Data Analysis {#sec-FDA}

**Firstly, due to the variable of Sex,Household_type,Electricity are
category variables,so we set these as factor using factor()**

```{r}
# deal with the category varibales
data$Sex<-factor(data$Sex)
data$Household_Type<-factor(data$Household_Type)
data$Electricity<-factor(data$Electricity)
```

```{r}
data$log_Income <- log(data$Income)
data$log_Food_Expenditure <- log(data$Food_Expenditure)
skim(data)
```

**Then, we fit a generalized linear model with a Poisson family and log
link between Family_members and explanatory variables:**

```{r}
# Fitting model1
#full-model
model1 <- glm(Family_members ~ Food_Expenditure +Income+Sex + Head_Age + 
                    Household_Type + Floor_Area + House_Age + 
                    Num_bedrooms + Electricity, family = poisson(link = "log"), data = data)
summary(model1)
```

**Fit Poisson regression model 1 between family members and explanatory
variables:**

$$
\log \left( \widehat{\text{Family Members}}_i \right) = \hat{\alpha} + \hat{\beta}_1 \cdot \text{Food Expenditure}_i + \hat{\beta}_2 \cdot \text{Income}_i + \hat{\beta}_3 \cdot \text{SexMale}_i + \hat{\beta}_4 \cdot \text{Head Age}_i + \hat{\beta}_5 \cdot \text{Household TypeSingle Family}_i + \hat{\beta}_6 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i + \hat{\beta}_7 \cdot \text{Floor Area}_i + \hat{\beta}_8 \cdot \text{House Age}_i + \hat{\beta}_9 \cdot \text{Num Bedrooms}_i + \hat{\beta}_{10} \cdot \text{Electricity}_i
$$

$$
= 1.597 + 2.930 \times 10^{-6} \cdot \text{Food Expenditure}_i - 2.385 \times 10^{-7} \cdot \text{Income}_i + 2.631 \times 10^{-1} \cdot \text{SexMale}_i - 3.797 \times 10^{-3} \cdot \text{Head Age}_i - 3.467 \cdot \text{Household TypeSingle Family}_i - 1.058 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i - 4.940 \times 10^{-2} \cdot \text{Floor Area}_i - 3.715 \times 10^{-3} \cdot \text{House Age}_i + 5.011 \times 10^{-2} \cdot \text{Num Bedrooms}_i - 9.028 \times 10^{-2} \cdot \text{Electricity}_i
$$

Where:

$\hat{\alpha} = 1.597$ is the intercept coefficient. It represents the
log of the expected number of Family Members when all explanatory
variables are zero. On the original scale, the expected number of family
members is $\exp(1.597) \approx 4.94$.

$\hat{\beta}_1 = 2.930 \times 10^{-6}$ is the coefficient for Food
Expenditure. For every one unit increase in food expenditure, the log of
the expected count of family members increases by
$2.930 \times 10^{-6}$.

The rate ratio is $\exp(2.930 \times 10^{-6}) \approx 1.00000293$, which
indicates a very small positive association.

$\hat{\beta}_2 = -2.385 \times 10^{-7}$ is the coefficient for Income.
For every one unit increase in income, the log of the expected family
members decreases by $2.385 \times 10^{-7}$.

The rate ratio is $\exp(-2.385 \times 10^{-7}) \approx 0.99999976$,
suggesting a very small negative effect.

$\hat{\beta}_3 = 2.631 \times 10^{-1}$ is the coefficient for SexMale.
Male-headed households are associated with an increase in the log of
expected family members by $0.2631$.

The rate ratio is $\exp(0.2631) \approx 1.301$, meaning male-headed
households are expected to have about 30.1% more family members than
female-headed households, holding other variables constant.

$\hat{\beta}_4 = -3.797 \times 10^{-3}$ is the coefficient for Head Age.
For each additional year of the household head's age, the log of
expected family members decreases by $0.003797$.

The rate ratio is $\exp(-0.003797) \approx 0.9962$, indicating that each
additional year of the head’s age is associated with a 0.38% decrease in
the expected number of family members.

$\hat{\beta}_5 = -3.467$ is the coefficient for Household TypeSingle
Family.

The rate ratio is $\exp(-3.467) \approx 0.031$, meaning that households
categorized as Single Family\` are expected to have about 96.9% fewer
family members compared to the reference group.

$\hat{\beta}_6 = -1.058$ is the coefficient for Household TypeTwo or
More Nonrelated Persons/Members.

The rate ratio is $\exp(-1.058) \approx 0.347$, meaning these households
have an expected family size 65.3% lower than the reference group.

$\hat{\beta}_7 = -4.940 \times 10^{-2}$ is the coefficient for Floor
Area. For every one unit increase in floor area, the log of expected
family members decreases by $0.0494$.

The rate ratio is $\exp(-0.0494) \approx 0.9518$, suggesting a 4.82%
decrease in family size for each unit increase in floor area.

$\hat{\beta}_8 = -3.715 \times 10^{-3}$ is the coefficient for House
Age. For each additional year of house age, the expected number of
family members decreases slightly.

The rate ratio is $\exp(-0.003715) \approx 0.9963$, or a 0.37% decrease
per additional year.

$\hat{\beta}_9 = 5.011 \times 10^{-2}$ is the coefficient for Num
Bedrooms. For each additional bedroom, the log of expected family
members increases by $0.05011$.

The rate ratio is $\exp(0.05011) \approx 1.0514$, which means each extra
bedroom is associated with a 5.14% increase in the expected number of
family members.

$\hat{\beta}_{10} = -9.028 \times 10^{-2}$ is the coefficient for
Electricity.

The rate ratio is $\exp(-0.09028) \approx 0.9137$, meaning that
households with electricity have an expected family size 8.63% lower
compared to those without electricity (assuming 1 = electricity, 0 = no
electricity).

This suggests that variables like Food Expenditure, SexMale, and Num
Bedrooms have a positive association with family size, indicating that
increases in these variables are associated with a larger number of
family members. On the other hand, variables like Income, Head Age,
Household Type, Floor Area, House Age, and Electricity have a negative
association, suggesting that higher values or presence of these factors
are linked to fewer family members.

**Looking at the P-value,the number of the Floor_Area is 0.1464 large
than the 0.05,so we delete it.**

```{r}
#Fitting model2
#Delete Floor_Area 
model2 <- glm(Family_members ~ Food_Expenditure +Income+Sex + Head_Age + 
                    Household_Type  + House_Age + 
                    Num_bedrooms + Electricity, family = poisson(link = "log"), data = data)
summary(model2)
```

**After we deleted Floor_Area, we fit Poisson regression model 2 between
family members and the rest of explanatory variables:**

$$
\log \left( \widehat{\text{Family Members}}_i \right) = \hat{\alpha} + \hat{\beta}_1 \cdot \text{Food Expenditure}_i + \hat{\beta}_2 \cdot \text{Income}_i + \hat{\beta}_3 \cdot \text{SexMale}_i + \hat{\beta}_4 \cdot \text{Head Age}_i + \hat{\beta}_5 \cdot \text{Household TypeSingle Family}_i + \hat{\beta}_6 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i + \hat{\beta}_7 \cdot \text{House Age}_i + \hat{\beta}_8 \cdot \text{Num Bedrooms}_i + \hat{\beta}_9 \cdot \text{Electricity}_i
$$

$$
= 1.596 + 2.935 \times 10^{-6} \cdot \text{Food Expenditure}_i - 2.532 \times 10^{-7} \cdot \text{Income}_i + 2.634 \times 10^{-1} \cdot \text{SexMale}_i - 3.852 \times 10^{-3} \cdot \text{Head Age}_i - 3.470 \times 10^{-1} \cdot \text{Household TypeSingle Family}_i - 1.019 \times 10^{-1} \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i - 3.760 \times 10^{-3} \cdot \text{House Age}_i + 4.445 \times 10^{-2} \cdot \text{Num Bedrooms}_i - 9.133 \times 10^{-2} \cdot \text{Electricity}_i
$$

Where:

$\hat{\alpha} = 1.596$ is the intercept coefficient, representing the
log of the expected number of Family Members when all explanatory
variables equal zero. On the original scale, the expected number of
family members is $\exp(1.596) \approx 4.94$.

$\hat{\beta}_1 = 2.935 \times 10^{-6}$ for Food Expenditure suggests a
very slight positive association. Each unit increase in food expenditure
increases the expected count of family members by approximately
$\exp(2.935 \times 10^{-6}) \approx 1.000002935$.

$\hat{\beta}_2 = -2.532 \times 10^{-7}$ for Income indicates a slight
negative association, where higher income slightly reduces the expected
family count by a factor of
$\exp(-2.532 \times 10^{-7}) \approx 0.999999747$.

$\hat{\beta}_3 = 2.634 \times 10^{-1}$ for SexMale shows male-headed
households have about $\exp(0.2634) \approx 1.3014$ times (approximately
30.14%) more expected family members than female-headed households.

$\hat{\beta}_4 = -3.852 \times 10^{-3}$ for Head Age indicates each
additional year of the household head’s age decreases the expected
family members by a factor of $\exp(-0.003852) \approx 0.9962$ (about a
0.38% reduction).

$\hat{\beta}_5 = -3.470 \times 10^{-1}$ for Household TypeSingle Family
implies such households have $\exp(-0.3470) \approx 0.7067$ (about
29.33% fewer) family members compared to the reference.

$\hat{\beta}_6 = -1.019 \times 10^{-1}$ for Household TypeTwo or More
Nonrelated Persons/Members shows these households have
$\exp(-0.1019) \approx 0.9031$ (around 9.69% fewer) family members
compared to the reference.

$\hat{\beta}_7 = -3.760 \times 10^{-3}$ for House Age means each
additional year of house age slightly reduces expected family members by
a factor of $\exp(-0.003760) \approx 0.9962$, approximately a 0.38%
decrease.

$\hat{\beta}_8 = 4.445 \times 10^{-2}$ for Num Bedrooms suggests each
additional bedroom increases expected family size by
$\exp(0.04445) \approx 1.0455$, or about 4.55%.

$\hat{\beta}_9 = -9.133 \times 10^{-2}$ for Electricity shows households
with electricity have an expected family size
$\exp(-0.09133) \approx 0.9127$, approximately 8.73% fewer than those
without electricity.

This model suggests that food expenditure, male-headed households, and
number of bedrooms positively influence family size. Conversely, higher
income, older household heads, household type, older houses, and
availability of electricity are negatively associated with family size.

```{r}
#Fitting model3
#Using the log transform
model3 <- glm(Family_members ~ log_Food_Expenditure +log_Income+Sex + Head_Age + 
                    Household_Type  + House_Age + 
                    Num_bedrooms + Electricity, family = poisson(link = "log"), data = data)
summary(model3)
```

**And then we use the log transform to fit Poisson regression model 3
between family members and the rest of explanatory variables:**

$$
\log \left( \widehat{\text{Family Members}}_i \right) = \hat{\alpha} + \hat{\beta}_1 \cdot \text{log Food Expenditure}_i + \hat{\beta}_2 \cdot \text{log Income}_i + \hat{\beta}_3 \cdot \text{SexMale}_i + \hat{\beta}_4 \cdot \text{Head Age}_i + \hat{\beta}_5 \cdot \text{Household TypeSingle Family}_i + \hat{\beta}_6 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i + \hat{\beta}_7 \cdot \text{House Age}_i + \hat{\beta}_8 \cdot \text{Num Bedrooms}_i + \hat{\beta}_9 \cdot \text{Electricity}_i
$$

$$
= -2.861 + 0.5764 \cdot \text{log Food Expenditure}_i - 0.1440 \cdot \text{log Income}_i + 0.2047 \cdot \text{SexMale}_i - 0.002736 \cdot \text{Head Age}_i - 0.2900 \cdot \text{Household TypeSingle Family}_i - 0.02888 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i - 0.003839 \cdot \text{House Age}_i + 0.01466 \cdot \text{Num Bedrooms}_i - 0.1582 \cdot \text{Electricity}_i
$$

Where:

$\hat{\alpha} = -2.861$ is the intercept coefficient. It represents the
log of the expected number of Family Members when all explanatory
variables are zero. On the original scale, the expected number of family
members is $\exp(-2.861) \approx 0.057$.

$\hat{\beta}_1 = 0.5764$ for log Food Expenditure indicates a strong
positive association. Each one-unit increase in log food expenditure
increases the expected count of family members by a factor of
$\exp(0.5764) \approx 1.779$, or approximately 77.9%.

$\hat{\beta}_2 = -0.1440$ for log Income suggests a negative
association. Each one-unit increase in log income reduces the expected
count of family members by a factor of $\exp(-0.1440) \approx 0.866$, or
about 13.4% fewer family members.

$\hat{\beta}_3 = 0.2047$ for SexMale means male-headed households are
expected to have $\exp(0.2047) \approx 1.227$ times more family members
than female-headed households, an increase of about 22.7%.

$\hat{\beta}_4 = -0.002736$ for Head Age implies each additional year in
the household head’s age reduces the expected number of family members
by a factor of $\exp(-0.002736) \approx 0.9973$, approximately a 0.27%
decrease.

$\hat{\beta}_5 = -0.2900$ for Household TypeSingle Family shows these
households have $\exp(-0.2900) \approx 0.748$, or 25.2% fewer family
members than the reference household type.

$\hat{\beta}_6 = -0.02888$ for Household TypeTwo or More Nonrelated
Persons/Members shows a negligible effect, with a factor of
$\exp(-0.02888) \approx 0.9715$, or a 2.85% decrease (not statistically
significant as p \> 0.05).

$\hat{\beta}_7 = -0.003839$ for House Age indicates each additional year
of house age reduces expected family members by a factor of
$\exp(-0.003839) \approx 0.9962$, about a 0.38% decrease.

$\hat{\beta}_8 = 0.01466$ for Num Bedrooms suggests each additional
bedroom increases expected family size by
$\exp(0.01466) \approx 1.0147$, or about 1.47% (not statistically
significant as p \> 0.05).

$\hat{\beta}_9 = -0.1582$ for Electricity shows households with
electricity have an expected family size $\exp(-0.1582) \approx 0.8538$,
or approximately 14.62% fewer family members than those without
electricity.

This model suggests that log food expenditure, male-headed households,
and number of bedrooms positively influence family size. Conversely,
higher log income, older household heads, household type, older houses,
and having electricity are negatively associated with family size.

```{r}
#Fitting model4
#Delete number of bedrooms
model4 <- glm(Family_members ~ log_Food_Expenditure +log_Income+Electricity+Sex + Head_Age + 
                    Household_Type  + House_Age 
                    , family = poisson(link = "log"), data = data)
summary(model4)
```

**After we deleted number of bedrooms, we fit Poisson regression model 4
between family members and the rest of explanatory variables:**

$$
\log \left( \widehat{\text{Family Members}}_i \right) = \hat{\alpha} + \hat{\beta}_1 \cdot \text{log Food Expenditure}_i + \hat{\beta}_2 \cdot \text{log Income}_i + \hat{\beta}_3 \cdot \text{Electricity}_i + \hat{\beta}_4 \cdot \text{SexMale}_i + \hat{\beta}_5 \cdot \text{Head Age}_i + \hat{\beta}_6 \cdot \text{Household TypeSingle Family}_i + \hat{\beta}_7 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i + \hat{\beta}_8 \cdot \text{House Age}_i
$$

$$
= -2.966 + 0.5802 \cdot \text{log Food Expenditure}_i - 0.1369 \cdot \text{log Income}_i - 0.1567 \cdot \text{Electricity}_i + 0.2030 \cdot \text{SexMale}_i - 0.002665 \cdot \text{Head Age}_i - 0.2908 \cdot \text{Household TypeSingle Family}_i - 0.03024 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i - 0.003698 \cdot \text{House Age}_i
$$

Where:

$\hat{\alpha} = -2.966$ is the intercept coefficient. It represents the
log of the expected number of Family Members when all explanatory
variables are zero. On the original scale, the expected number of family
members is $\exp(-2.966) \approx 0.051$.

$\hat{\beta}_1 = 0.5802$ for log Food Expenditure indicates a strong
positive association. Each one-unit increase in log food expenditure
increases the expected count of family members by a factor of
$\exp(0.5802) \approx 1.786$, or approximately 78.6%.

$\hat{\beta}_2 = -0.1369$ for log Income suggests a negative
association. Each one-unit increase in log income reduces the expected
count of family members by a factor of $\exp(-0.1369) \approx 0.872$, or
about 12.8% fewer family members.

$\hat{\beta}_3 = -0.1567$ for Electricity shows households with
electricity have an expected family size $\exp(-0.1567) \approx 0.855$,
approximately 14.5% fewer family members than those without electricity.

$\hat{\beta}_4 = 0.2030$ for SexMale means male-headed households are
expected to have $\exp(0.2030) \approx 1.225$ times more family members
than female-headed households, an increase of about 22.5%.

$\hat{\beta}_5 = -0.002665$ for Head Age implies each additional year in
the household head’s age reduces the expected number of family members
by a factor of $\exp(-0.002665) \approx 0.9973$, approximately a 0.27%
decrease.

$\hat{\beta}_6 = -0.2908$ for Household TypeSingle Family shows these
households have $\exp(-0.2908) \approx 0.748$, or 25.2% fewer family
members than the reference household type.

$\hat{\beta}_7 = -0.03024$ for Household TypeTwo or More Nonrelated
Persons/Members shows a negligible effect, with a factor of
$\exp(-0.03024) \approx 0.9702$, or a 2.98% decrease (not statistically
significant as p \> 0.05).

$\hat{\beta}_8 = -0.003698$ for House Age indicates each additional year
of house age reduces expected family members by a factor of
$\exp(-0.003698) \approx 0.9963$, about a 0.37% decrease.

This model suggests that log food expenditure, male-headed households
positively influence family size. Conversely, higher log income, older
household heads, household type, older houses, and having electricity
are negatively associated with family size.

# Selecting models

**In the previous analysis, variable selection was based on p-values to
identify statistically significant predictors. In this section, model
selection was conducted using AIC and BIC as criteria to compare overall
model performance.**

```{r}
models <- list(model1, model2, model3, model4)
select_table <- data.frame(
  Model = paste0("model", 1:4),
  AIC = sapply(models, AIC)
)
print(select_table)
```

**According to the table ,we find that model4 has the lowest AIC ,so we
decide to choose the model 4 as our final model.**

```{r}
coef(model4)
```

The coefficients from `model4` represent the estimated changes in the
log of the expected number of family members for each unit increase in
the predictor variables. Positive coefficients indicate an increase in
the log expected count, while negative coefficients indicate a decrease.
For example, `log_Food_Expenditure` has a coefficient of 0.5802,
suggesting that as food expenditure increases, the expected log count of
family members also increases. Conversely, `log_Income` has a negative
coefficient of -0.1369, indicating a decrease in the expected log count
with higher income.

```{r}
exp(coef(model4))
```

The exponentiated coefficients represent the multiplicative effect on
the expected number of family members. A value greater than 1 indicates
an increase, while a value less than 1 indicates a decrease. For
instance, the exponentiated coefficient for `log_Food_Expenditure` is
1.7863, meaning a one-unit increase in food expenditure is associated
with a 78.6% increase in expected family members. On the other hand,
`Electricity1` has a value of 0.8549, indicating households with
electricity are expected to have approximately 14.5% fewer family
members compared to those without electricity.

# Fitted values and residuals

```{r}
#| fig-cap: "Figure 5: Diagnostic Plots for Model 4"

resp <- resid(model4, type = "pearson")
resd <- resid(model4, type = "deviance")
p1<- ggplot(model4, aes(sample = resp)) + geom_point(stat = "qq", color = "#7fc97f") + 
  ylab("Pearson residuals")
p2<- ggplot(model4, aes(sample = resd)) + geom_point(stat = "qq", color = "#7fc97f") + 
  ylab("Deviance residuals") 
p3<- ggplot(model4, aes(x = predict(model1, type="link"), y =resd))+ 
  geom_point(col = "#7fc97f") + 
  ylab("Deviance residuals") + xlab("Linear predictor")
grid.arrange(p1, p2, p3, nrow = 1)
```

Figure 5 presents diagnostic plots for assessing the goodness-of-fit and
assumptions of Model 4.

-   The **first plot** is a Q-Q plot of Pearson residuals, which shows
    some deviations from the theoretical line, especially at the tails,
    suggesting potential departures from normality.

-   The **second plot** is a Q-Q plot of deviance residuals, displaying
    a similar pattern with slight deviations at the extremes.

-   The **third plot** plots deviance residuals against the linear
    predictor. The residuals appear randomly scattered without a clear
    pattern, suggesting no major issues with heteroscedasticity, though
    a slight funnel shape may indicate some variance differences at
    different predictor values.

These diagnostic plots suggest that Model 4 fits the data reasonably
well, but some minor deviations in the residuals indicate areas for
potential improvement.

```{r}
#| fig-cap: "Figure 6: Plot of Squared Residuals Versus Predicted Values (Model 4)"

ggplot(model4, aes(x=log(fitted(model4)), y=log((data$Family_members-fitted(model4))^2)))+
  geom_point(col="#f46d43") +
  geom_abline(slope=1, intercept=0, col="#a6d96a", size=1) +
  ylab(expression((y-hat(mu))^2)) + xlab(expression(hat(mu)))

```

```{r}
overdispersion_test <- sum(residuals(model4, type = "pearson")^2) / df.residual(model4)
overdispersion_test
```

Figure 6 presents the plot of squared residuals $(y - \hat{y})^2$
against the predicted values $\hat{y}$ from Model 4. This plot is used
to assess the variance structure and check for potential
heteroscedasticity in the model.

The dispersion of residuals around the horizontal axis suggests some
variation in the spread of residuals at different levels of predicted
values. The green line represents the trend in variance, and its slight
upward slope indicates a modest increase in variance as the predicted
values increase. However, the pattern appears relatively contained, and
the dispersion parameter of 0.6354 (shown below the plot) indicates that
the model accounts for overdispersion reasonably well.

This plot suggests that Model 4 fits the data adequately, with no strong
evidence of severe heteroscedasticity.

# Using Quasi-poisson model

```{r}
model4_quasi <- glm(Family_members ~ log_Food_Expenditure +log_Income+Electricity + Head_Age + 
                    Household_Type  + House_Age + Sex
                    , family = quasipoisson(link = "log"), data = data)
summary(model4_quasi)
```

**Fit Quasi-Poisson regression model between family members and
explanatory variables:**

$$
\log \left( \widehat{\text{Family Members}}_i \right) = \hat{\alpha} + \hat{\beta}_1 \cdot \text{log Food Expenditure}_i + \hat{\beta}_2 \cdot \text{log Income}_i + \hat{\beta}_3 \cdot \text{Electricity}_i + \hat{\beta}_4 \cdot \text{Head Age}_i + \hat{\beta}_5 \cdot \text{Household TypeSingle Family}_i + \hat{\beta}_6 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i + \hat{\beta}_7 \cdot \text{House Age}_i + \hat{\beta}_8 \cdot \text{SexMale}_i
$$

$$
= -2.9657 + 0.5802 \cdot \text{log Food Expenditure}_i - 0.1369 \cdot \text{log Income}_i - 0.1567 \cdot \text{Electricity}_i - 0.002665 \cdot \text{Head Age}_i - 0.2907 \cdot \text{Household TypeSingle Family}_i - 0.03024 \cdot \text{Household TypeTwo or More Nonrelated Persons/Members}_i - 0.003698 \cdot \text{House Age}_i + 0.2030 \cdot \text{SexMale}_i
$$

Where:

$\hat{\alpha} = -2.9657$ is the intercept coefficient. It represents the
log of the expected number of Family Members when all explanatory
variables are zero. On the original scale, the expected number of family
members is $\exp(-2.9657) \approx 0.0514$.

$\hat{\beta}_1 = 0.5802$ for log Food Expenditure indicates a strong
positive association. Each one-unit increase in log food expenditure
increases the expected count of family members by a factor of
$\exp(0.5802) \approx 1.786$, or approximately 78.6%.

$\hat{\beta}_2 = -0.1369$ for log Income suggests a negative
association. Each one-unit increase in log income reduces the expected
count of family members by a factor of $\exp(-0.1369) \approx 0.872$, or
about 12.8% fewer family members.

$\hat{\beta}_3 = -0.1567$ for Electricity shows households with
electricity have an expected family size $\exp(-0.1567) \approx 0.855$,
approximately 14.5% fewer family members than those without electricity.

$\hat{\beta}_4 = -0.002665$ for Head Age implies each additional year in
the household head’s age reduces the expected number of family members
by a factor of $\exp(-0.002665) \approx 0.9973$, approximately a 0.27%
decrease.

$\hat{\beta}_5 = -0.2907$ for Household TypeSingle Family shows these
households have $\exp(-0.2907) \approx 0.748$, or 25.2% fewer family
members than the reference household type.

$\hat{\beta}_6 = -0.03024$ for Household TypeTwo or More Nonrelated
Persons/Members shows a negligible effect, with a factor of
$\exp(-0.03024) \approx 0.9702$, or a 2.98% decrease (not statistically
significant as p \> 0.05).

$\hat{\beta}_7 = -0.003698$ for House Age indicates each additional year
of house age reduces expected family members by a factor of
$\exp(-0.003698) \approx 0.9963$, about a 0.37% decrease.

$\hat{\beta}_8 = 0.2030$ for SexMale means male-headed households are
expected to have $\exp(0.2030) \approx 1.225$ times more family members
than female-headed households, an increase of about 22.5%.

This Quasi-Poisson model suggests that log food expenditure and
male-headed households positively influence family size. Conversely,
higher log income, older household heads, household type, older houses,
and having electricity are negatively associated with family size.

```{r}
coef(model4_quasi)
```

The coefficients from `model4_quasi` represent the changes in the log of
the expected number of family members for a one-unit increase in each
predictor, holding other variables constant. Positive coefficients
(e.g., `log_Food_Expenditure = 0.5802`) indicate an increase in the log
expected count, while negative coefficients (e.g.,
`log_Income = -0.1369`) suggest a decrease in the log expected count.

```{r}
exp(coef(model4_quasi))
```

The exponentiated coefficients reflect the multiplicative change (rate
ratio) in the expected number of family members for a one-unit increase
in each predictor.

-   `log_Food_Expenditure` has an exponentiated coefficient of 1.7863,
    meaning each unit increase in log food expenditure is associated
    with a 78.6% increase in the expected family size.

-   `Electricity1` has an exponentiated coefficient of 0.8549,
    indicating that households with electricity have approximately 14.5%
    fewer expected family members compared to households without
    electricity.

# Conclusions {#sec-Conc}

## Conclusions

This project investigated the household-related factors influencing the
number of people living in a household, using data from the Family
Income and Expenditure Survey (FIES) in the Philippines. The analysis
was conducted through a series of Generalised Linear Models (GLMs), with
a focus on variable selection and model performance evaluation.

The analysis demonstrates that the most influential household-related
variables on the number of people living in a household are food
expenditure, income, electricity access, household head's sex and age,
household type, and house age. These factors significantly affect
household size, providing valuable insights for government
decision-making in resource allocation and policy development.

## Key Findings

Initial variable selection was based on p-values, identifying key
predictors such as food expenditure, household income, head age, sex of
the household head, household type, house age, and access to
electricity. Model 4, selected for its lowest AIC and BIC values,
initially appeared to perform best. However, diagnostic checks revealed
evidence of overdispersion, suggesting the Poisson model assumptions
were violated.

To address overdispersion, a Quasi-Poisson model was fitted. This model
adjusted for the dispersion parameter (0.6354), resulting in more
reliable estimates and inference. The final results indicate:

-   Positive influences on household size: higher food expenditure and
    male-headed households.

-   Negative influences: higher income, access to electricity, older
    household heads, single-family household types, and older houses.

-   No significant influence was found from households categorized as
    "Two or More Non-related Persons/Members."

## Limitations

-   The Quasi-Poisson model cannot provide AIC or BIC values, limiting
    its use for direct comparison with alternative models.

-   Potential issues such as zero inflation or more complex data
    distributions were not addressed in this analysis, which may affect
    model accuracy.

-   Interaction effects between explanatory variables were not
    considered. Including interaction terms may help capture more
    complex relationships and potentially improve model performance.

## Further Work

-   Explore models that can handle zero-inflated data or overdispersion
    more effectively, such as zero-inflated Poisson or negative binomial
    regression.

-   Investigate potential interaction effects between key explanatory
    variables to better understand their combined influence on household
    size.

-   Consider hierarchical or multilevel modeling approaches to account
    for regional variation or clustering within the data.

-   Conduct sensitivity analyses or validation using datasets from
    different time periods to assess the generalizability of the model
    findings.
